---
title: "Effects of Vitamin C Doses and Delivery Methods on Guinea Pig Tooth Growth"
output:
  html_document:
    keep_md: yes
---

## Synopsis

This analysis explores tbe impact of different vitamin C doses and delivery methods on the length of guinea pig teeth.

## Load R packages

Let's use the ToothGrowth dataset and load the necessary packages

```{r}
attach(ToothGrowth)
data <- ToothGrowth
library(ggplot2)
library(dplyr)
```

## Exploratory Analysis and Visual Inspection

First, explore the tooth lengths by looking at vitamin doses (either 0.5mg, 1mg or 2mg) or vitamin delivery method (orange juice (OJ) or ascorbic acid(VC))

```{r}
ggplot(data = data, aes(x = len)) + geom_histogram(binwidth = 0.5, fill = "salmon") + facet_grid(dose ~ supp) + ggtitle("Length of guinea pig tooth by vitamin dose and vitamin delivery method") + xlab("tooth length") + ylab("number of guinea pigs")
```

Visually, it is pretty clear that when doses of vitamin C is higher, the tooth length is also higher. This is true for both delivery methods.

## Confirm with Inferential Analysis

Let's confirm with t-test. Create subsets of data to compare OJ & VC and to compare 0.5mg and 2.0mg doses.

```{r}
dose0.5 <- filter(data, dose == 0.5)
dose2.0 <- filter(data, dose == 2.0)
oj <- filter(data, supp == "OJ")
vc <- filter(data, supp == "VC")
```

Set x as tooth lengths as a result of 0.5mg dose and y as tooth lengths as a result of 2.0mg. Let's perform a t-test.

```{r}
t.test(dose0.5$len, dose2.0$len)
```

Noticed that the mean tooth length for guinea pigs is 26.1 for the 2.0mg dose compared 10.6 for the 0.5mg dose.

Let's compare the OJ and VC groups:

```{r}
t.test(oj$len, vc$len)
```

This confirms that guinea pigs tend to have longer tooth on average when given orange juice compared to ascorbic acid.